NeuroElectro: A Window to the World's Neuron Electrophysiology Data

### 2014-03-17 neuroelectro_frontiers/main.tex
Added tex line breaks for easy diffability
Revisions for reviewers

\title[NeuroElectro.org: text-mining neurophysiology data]{NeuroElectro: A Window to the World's Neurophysiology Data}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}

@@

We obtained tens of thousands of potentially relevant full article texts directly from publisher websites.  We first identified potential articles that were likely to contain information relevant to neuron biophysics using the native search functions provided within the journal websites and only downloaded articles containing a specific list of terms including ``input resistance'' and ``resting membrane potential'' (Fig. \ref{text_mining_workflow}).  This pre-selection step allowed us to identify and download only articles that contained data relevant to our project.  Upon identifying candidate articles, we then downloaded the full text of each potentially-relevant article as HTML; articles downloaded from the publisher Elsevier (e.g., \emph{Neuron and Brain Research}) were downloaded as XML using the provided text-mining API and subsequently converted to HTML.  We chose to work with HTML (as opposed to PDF or XML) because HTML provides a machine-readable markup of the article's content, allowing us to easily identify relevant elements within the article -- such as data tables and the methods section -- using publicly available HTML-parsing tools (here we used the Beautiful Soup HTML-processing library implemented in Python: \url{http://www.crummy.com/software/BeautifulSoup/bs4/doc/}).  Furthermore, because HTML is a single semi-structured standard used across publishers, we could write relatively generic HTML-processing algorithms applicable to content published across journals.  However, our focus on using HTML limits us to relatively newer articles -- typically those published after 1997 -- because most publications before this time are only available as scanned PDF files.

@@

@@

>>>>

\title[NeuroElectro.org: text-mining neurophysiology data]{NeuroElectro: A Window to the World's Neuron Electrophysiology Data}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}

@@

We obtained tens of thousands of potentially relevant full article texts directly from publisher websites.  
We first identified potential articles that were likely to contain information relevant to neuron biophysics using the native search functions provided within the journal websites and only downloaded articles containing a specific list of terms including ``input resistance'' and ``resting membrane potential'' (Fig. \ref{text_mining_workflow}).  
This pre-selection step allowed us to identify and download only articles that contained data relevant to our project.  
Upon identifying candidate articles, we then downloaded the full text of each potentially-relevant article as HTML; 
articles downloaded from the publisher Elsevier (e.g., \emph{Neuron and Brain Research}) were downloaded as XML using the provided text-mining API and subsequently converted to HTML.  
We chose to work with HTML (as opposed to PDF or XML) because HTML provides a machine-readable markup of the article's content, allowing us to easily identify relevant elements within the article -- such as data tables and the methods section -- using publicly available HTML-parsing tools (here we used the Beautiful Soup HTML-processing library implemented in Python: \url{http://www.crummy.com/software/BeautifulSoup/bs4/doc/}).  
Furthermore, because HTML is a single semi-structured standard used across publishers, we could write relatively generic HTML-processing algorithms applicable to content published across journals.  
Our focus on using HTML limits us to relatively newer articles - typically those published after 1996 - because before this time most publications are only available as scanned PDF files.  
However, because the rate of publication across the field has grown exponentially, this HTML-available subset constitutes the vast majority of neuroscience articles ever published.

@@

\subsection{Preliminary use of NeuroElectro in scientific work}
The NeuroElectro project is intended to facilitate scientific investigation by providing easy access to large quantities of data about neurons.  
Because the data is machine-readable, we have already begun to conduct several analyses that would not be possible without this resource.  
First, we have begun an investigation of the relationships between neurons defined by the similarity of their electrophysiological properties.  
This information can be used to make predictions about as yet unmeasured properties.  
Second, we have begun to explore the relationship between patterns of gene expression (using both the Allen Brain Atlas and single cell qPCR approaches) and electrophysiological properties of neurons.  
Third, we have begun automated testing of quantitative neuron models, under the reasonable assumption that these models should be constrained by the available experimental data.  
These projects are described in manuscripts currently in preparation.

@@

Future developments in machine extraction of data from the scientific literature will be of great benefit.  
These should include better semantic understanding of context, ranging from relatively unambiguous notations such as units, to syntax-parsing that relates objects of study to their reported properties in free-form prose.  
Much progress has been made by computer scientists in some of these areas, and more future engagement with their research should enable vastly more data to be extracted from the literature.  
In the mean time, experimentalists who would like their data to be easily curated should strongly consider using data tables in future publications.
